Greenhouse gas emissions model - making scenarios

### 2021-12-20 Scenarios_report.Rmd
added herd structure (fewer) males scenarios and cleaned up a bit

```{r, warning=FALSE, include=FALSE}
#install and load packages
library(DiagrammeR)
# devtools::install_github("eikeluedeling/decisionSupport")
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
# devtools::install_github("CWWhitney/uncertainty")
library(uncertainty)

@@

```{r monte_carlo_simulation, warning=FALSE}

@@

@@

# Normalizing milk yield and emissions per head

```{r}

animals_per_run<-rowSums(GHG_simulation_scenarios$x[paste0("N_animal_type_",1:12)])

milk_yield_per_head<-GHG_simulation_scenarios$y$milk_yield/animals_per_run
total_emissions_per_head<-GHG_simulation_scenarios$y$on_farm/animals_per_run

@@

```{r name_in_out_vars}
in_var  <- milk_yield_per_head # "Total Milk Yield"
out_var <- total_emissions_per_head # "Total emissions"
```

@@

```{r uncertainty_package}
uncertainty::varscatter(in_var = in_var,
                        out_var = out_var,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
```
```{r varkernel}
uncertainty::varkernel(in_var = in_var, 
                       out_var = out_var,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
```

@@

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = in_var, 
                                 out_var = out_var, 
                                 min_in_var = round(min(in_var)),
                                 max_in_var = round(max(in_var)), 
                                 xlab_vars = "Total emissions given milk yield range")
```

@@

Here's a wild idea: Can we simply change values in the GHG_simulation_scenarios$x table now and apply the model to the resulting table? We may not have to draw new values (and that actually wouldn't even be good, because this would dilute the intervention effect).

@@

```{r}
x_breed_intervention<-GHG_simulation_scenarios$x

adoption_rate<-0.6

for(i in 1:nrow(x_breed_intervention))
{if(rbinom(1,1,adoption_rate))
    x_breed_intervention[i,colnames(breed_intervention)[2:ncol(breed_intervention)]]<-
      x_breed_intervention[i,colnames(breed_intervention)[2:ncol(breed_intervention)]] * 
      breed_intervention[which(paste0("Farm_",breed_intervention$Farm)==x_breed_intervention$Scenario[i]),
                         2:ncol(breed_intervention)]
  # implement changes in x file
}
  breed_res<-run_model(x_breed_intervention,ghg_emissions)

  animals_per_run<-rowSums(breed_res$x[paste0("N_animal_type_",1:12)])

breed_milk_yield_per_head<-breed_res$y$milk_yield/animals_per_run
breed_total_emissions_per_head<-breed_res$y$on_farm/animals_per_run

@@

@@

uncertainty::varscatter(in_var = breed_milk_yield_per_head,
                        out_var = breed_total_emissions_per_head,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run,
                        out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = breed_milk_yield_per_head, 
                       out_var = breed_total_emissions_per_head,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run, 
                       out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernelslicerange(in_var = breed_res$y$milk_yield/animals_per_run, 
                                 out_var = breed_res$y$on_farm/animals_per_run, 
                                 min_in_var = round(min(breed_res$y$milk_yield/animals_per_run)),
                                 max_in_var = round(max(breed_res$y$milk_yield/animals_per_run)), 
                                 xlab_vars = "Total emissions given milk yield range")

uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$milk_yield/animals_per_run, 
                                 out_var = GHG_simulation_scenarios$y$on_farm/animals_per_run, 
                                 min_in_var = round(min(GHG_simulation_scenarios$y$milk_yield/animals_per_run)),
                                 max_in_var = round(max(GHG_simulation_scenarios$y$milk_yield/animals_per_run)), 
                                 xlab_vars = "Total emissions given milk yield range")

@@

```{r}
breed_change_in_CO2_eq_per_animal<-breed_res$y$on_farm/animals_per_run-
  GHG_simulation_scenarios$y$on_farm/animals_per_run

breed_change_in_milk_yield_per_animal<-breed_res$y$milk_yield/animals_per_run-
  GHG_simulation_scenarios$y$milk_yield/animals_per_run
```

@@

>>>>

```{r, warning=FALSE, include=FALSE}
#install and load packages
library(DiagrammeR)
# devtools::install_github("eikeluedeling/decisionSupport")
library(decisionSupport)
library(ggplot2)
library(kableExtra)
library(magrittr)
library(plyr)
library(rmarkdown)
# devtools::install_github("CWWhitney/uncertainty")
library(uncertainty)
library(chillR)

@@

```{r monte_carlo_simulation, warning=FALSE, eval=FALSE}

@@

```{r, eval=FALSE, include=FALSE}
dir.create("results")
save_temperature_scenarios(GHG_simulation_scenarios,"results","Baseline")
```

```{r, include=FALSE}
GHG_simulation_scenarios<-load_temperature_scenarios("results","Baseline")
class(GHG_simulation_scenarios)<-"mcSimulation"
```

@@

@@

@@

```{r uncertainty_package}
uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
```
```{r varkernel}
uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))
```

@@

```{r varkernelslicerange}
uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                                 out_var = GHG_simulation_scenarios$y$pc_on_farm, 
                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
                                 xlab_vars = "Total emissions given milk yield range")
```

@@

@@

```{r, eval=FALSE}
x_breed_intervention <- GHG_simulation_scenarios$x

adoption_rate <- 0.6

for (i in 1:nrow(x_breed_intervention))
{
  if (rbinom(1, 1, adoption_rate)) # if farmers adopt, make changes in x file
    x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] <-
      x_breed_intervention[i, colnames(breed_intervention)[2:ncol(breed_intervention)]] *
      breed_intervention[which(paste0("Farm_", breed_intervention$Farm) ==
                                 x_breed_intervention$Scenario[i]),
                         2:ncol(breed_intervention)]
}
# run model for modified x file
breed_res <- run_model(x_breed_intervention, ghg_emissions)

@@

```{r, include=FALSE, eval=FALSE}
breed_res$call<-"multi-scenario call"
save_temperature_scenarios(breed_res,"results","Breed_intervention")
```

```{r, include=FALSE}
adoption_rate <- 0.6
breed_res<-load_temperature_scenarios("results","Breed_intervention")
class(breed_res)<-"mcSimulation"
```

@@

uncertainty::varscatter(in_var = breed_res$y$pc_milk_yield,
                        out_var = breed_res$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = breed_res$y$pc_milk_yield,
                        out_var = breed_res$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernelslicerange(in_var = breed_res$y$pc_milk_yield,
                                 out_var = breed_res$y$pc_on_farm, 
                                 min_in_var = round(min(breed_res$y$pc_milk_yield)),
                                 max_in_var = round(max(breed_res$y$pc_milk_yield)), 
                                 xlab_vars = "Total emissions given milk yield range")

uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
                                 xlab_vars = "Total emissions given milk yield range")

@@

```{r}
breed_change_in_CO2_eq_per_animal<-breed_res$y$pc_on_farm-
  GHG_simulation_scenarios$y$pc_on_farm

breed_change_in_milk_yield_per_animal<-breed_res$y$pc_milk_yield-
  GHG_simulation_scenarios$y$pc_milk_yield
```

@@

### Retiring unproductive bulls

To simulate this scenario, we simply have to remove all animals of type 1 from the herds.

```{r , warning=FALSE, eval=FALSE}
x_nobull_intervention <- GHG_simulation_scenarios$x

nobull_adoption_rate <- 1.0

for (i in 1:nrow(x_nobull_intervention))
{
  if (rbinom(1, 1, nobull_adoption_rate)) # if farmers adopt, make changes in x file
    x_nobull_intervention$N_animal_type_1[i]<-0
}
# run model for modified x file
nobull_res <- run_model(x_nobull_intervention, ghg_emissions)

```{r, include=FALSE, eval=FALSE}
nobull_res$call<-"multi-scenario call"
save_temperature_scenarios(nobull_res,"results","noBull_intervention")
```

```{r, include=FALSE}
nobull_res<-load_temperature_scenarios("results","noBull_intervention")
class(nobull_res)<-"mcSimulation"
```

#### Evaluate impact of "Retiring unproductive bulls" intervention

We're assuming that all bulls >36 months old have been removed.

```{r, fig.show="hold", out.width="50%", fig.cap="No bulls intervention (left) vs. baseline (right)", message=FALSE}

uncertainty::varscatter(in_var = nobull_res$y$pc_milk_yield,
                        out_var = nobull_res$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = nobull_res$y$pc_milk_yield,
                        out_var = nobull_res$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernelslicerange(in_var = nobull_res$y$pc_milk_yield,
                                 out_var = nobull_res$y$pc_on_farm, 
                                 min_in_var = round(min(nobull_res$y$pc_milk_yield,na.rm=TRUE)),
                                 max_in_var = round(max(nobull_res$y$pc_milk_yield,na.rm=TRUE)), 
                                 xlab_vars = "Total emissions given milk yield range")

uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
                                 xlab_vars = "Total emissions given milk yield range")

### Reducing unproductive replacement males

The idea is to raise the off-take rate of replacement males by 23% (not sure why this number). We'll implement this by eliminating replacement males on all adopting farms, with a 23% adoption rate. The result should be quite similar.

```{r , warning=FALSE, eval=FALSE}
x_lessReplace_intervention <- GHG_simulation_scenarios$x

lessReplace_adoption_rate <- 0.23

for (i in 1:nrow(x_lessReplace_intervention))
{
  if (rbinom(1, 1, lessReplace_adoption_rate)) # if farmers adopt, make changes in x file
    x_lessReplace_intervention$N_animal_type_3[i]<-0
}
# run model for modified x file
lessReplace_res <- run_model(x_lessReplace_intervention, ghg_emissions)

```{r, include=FALSE, eval=FALSE}
lessReplace_res$call<-"multi-scenario call"
save_temperature_scenarios(lessReplace_res,"results","lessReplace_intervention")
```

```{r, include=FALSE}
lessReplace_res<-load_temperature_scenarios("results","lessReplace_intervention")
class(lessReplace_res)<-"mcSimulation"
```

#### Evaluate impact of "Retiring unproductive bulls" intervention

We're assuming that all bulls >36 months old have been removed.

```{r, fig.show="hold", out.width="50%", fig.cap="Fewer replacement males intervention (left) vs. baseline (right)", message=FALSE}

uncertainty::varscatter(in_var = lessReplace_res$y$pc_milk_yield,
                        out_var = lessReplace_res$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varscatter(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = lessReplace_res$y$pc_milk_yield,
                        out_var = lessReplace_res$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernel(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                        xlab = expression(Total~milk~yield~(kg~head^{-1}~year^{-1})),
                        ylab = expression(Total~emissions~(kg~CO[2]-eq~head^{-1}~year^{-1})))

uncertainty::varkernelslicerange(in_var = lessReplace_res$y$pc_milk_yield,
                                 out_var = lessReplace_res$y$pc_on_farm, 
                                 min_in_var = round(min(lessReplace_res$y$pc_milk_yield,na.rm=TRUE)),
                                 max_in_var = round(max(lessReplace_res$y$pc_milk_yield,na.rm=TRUE)), 
                                 xlab_vars = "Total emissions given milk yield range")

uncertainty::varkernelslicerange(in_var = GHG_simulation_scenarios$y$pc_milk_yield,
                        out_var = GHG_simulation_scenarios$y$pc_on_farm,
                                 min_in_var = round(min(GHG_simulation_scenarios$y$pc_milk_yield)),
                                 max_in_var = round(max(GHG_simulation_scenarios$y$pc_milk_yield)), 
                                 xlab_vars = "Total emissions given milk yield range")
